---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Unzipping and loading the data, identifying the NA values.
We also parse the dates.

```{r}
unzip("activity.zip")
data<-read.csv("activity.csv", na.strings="NA", stringsAsFactor=FALSE)
data$date <- as.Date(strptime(data$date, "%Y-%m-%d"))
```
Here is the data structure:
```{r echo=TRUE}
str(data)
```
The data now looks like:
```{r echo=TRUE}
head(data)
```

## What is mean total number of steps taken per day?

Histogram of the total number of steps taken each day.
Calculate and report the mean and median total number of steps taken per day.

```{r}
data.byday<-aggregate(data$steps,by=list(data$date),sum, na.rm=TRUE)
colnames(data.byday)<- c("date","steps")
hist(data.byday$steps)
```

The mean number of steps per day is given by:
```{r}
mean(data.byday$steps, na.rm=TRUE)
```

The median number of steps per day is given by:

```{r}
median(data.byday$steps, na.rm=TRUE)
```



## What is the average daily activity pattern?

Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis):
```{r}
data.byinterval<-aggregate(data$steps,by=list(data$interval),mean, na.rm=TRUE)
colnames(data.byinterval)<- c("interval","steps")
plot(data.byinterval$interval,data.byinterval$steps, type="l")
```

The 5-minute interval, on average across all the days in the dataset, containing the maximum number of steps is:
```{r}
data.byinterval$interval[which.max(data.byinterval$steps)]
```

## Imputing missing values
The total number of missing values in the dataset (i.e. the total number of rows with NAs) is:
```{r}
sum(!complete.cases(data))
```

Here is my strategy for filling in all of the missing values in the dataset.
Let's have a look as to where are the missing values:
```{r}
data$complete <- complete.cases(data)
aggregate(data$complete,by=list(data$date),sum, na.rm=TRUE)
```

All the missing values are located on specific dates. A give date has either no missing value (288 values), or all values are missing (0). I suggest we use the mean for that specific 5-minute interval across all days.

Let's create a new dataset that is equal to the original dataset but with the missing data filled in. Again, we use the average value for the interval when missing.

```{r}
library(dplyr)
data.complete<-merge(data, data.byinterval,by=c("interval"), all.x=TRUE, sort=FALSE)
data.complete$steps<-ifelse(data.complete$complete, 
                            data.complete$steps.x, data.complete$steps.y)
data.complete$steps.x<-NULL
data.complete$steps.y<-NULL
data.complete$complete<-NULL
data.complete<-arrange(data.complete, date)
```

The data now looks like:
```{r}
head(data.complete)
```

Let's create an histogram of the total number of steps taken each day, and compute the mean and median total number of steps taken per day.

```{r}
data.complete.byday<-aggregate(data$steps,by=list(data$date),sum, na.rm=TRUE)
colnames(data.complete.byday)<- c("date","steps")
hist(data.complete.byday$steps)
mean(data.complete.byday$steps, na.rm=TRUE)
median(data.complete.byday$steps, na.rm=TRUE)
```

The histogram, mean do not differ because we used mean values to impute the missing values. Median could differ in theory.

## Are there differences in activity patterns between weekdays and weekends?

Let's create a new factor variable in the dataset with two levels -- "weekday" and "weekend""
```{r}
data.complete$weekday<-ifelse((weekdays(data.complete$date)=="Samedi")| (weekdays(data.complete$date)=="Dimanche"),"weekend","weekday")
data.complete$weekday <- factor(data.complete$weekday)
```

Let's plot the average number of steps for each interval on weekend and weekdays:
```{r}
library(lattice)
data.complete.byinterval<-aggregate(data.complete$steps,by=list(data.complete$interval,data.complete$weekday),mean, na.rm=TRUE)
colnames(data.complete.byinterval)<- c("interval","weekday","steps")
xyplot(steps ~ interval|weekday , data = data.complete.byinterval, 
       layout=c(1,2),  type = "l")
```

We can notice that on weekends, the activity (steps) starts later in the day and stays at a higher level through the day.
